---
title: "MODIS data Santa Lucia extract"
author: "Willem Vervoort"
date: "`r format(Sys.Date(),'%d-%m-%Y')`"
output: 
    pdf_document:
      fig_width: 7
      fig_height: 6
      fig_caption: true
---

```{r setup, warning=F, message=F, echo=F}
# root dir
knitr::opts_knit$set(root.dir = 
                       "X:/PRJ-HPWC/SWAT_ETCalibration")
library(pander)
library(knitr)
```

# Introduction
This document covers downloading satellite data:

* MODIS16A ET (collection 5) using the package MODISTools

# Downloading MODIS ET data using MODISTools
The first component is a demonstration of using the package MODISTools in R to download the data. In this case the example will be based on extracting data for the midpoints of the subbasins of a SWAT model for the Santa Lucia catchment

## how to extract the midpoints  
This is fairly easy in ARCSWAT, the model actually generates the midpoints as part of the watershed delineation and you can access this through the table of data of the shape file.

## Downloading the data
Below is the actual download of the MODIS data, this requires an input file with the coordinates of the midpoints. 

```{r download}
# Install package if needed
if(!require(MODISTools)) install.packages("MODISTools")
library(MODISTools)

# read in file with xy locations (or create a data.frame)
xy.loc <- read.csv("Inputdata/subbasins_SantaLucia.csv")
# Following the MODIS tools manual
coords <- data.frame(lat=xy.loc$Lat, 
                     long=xy.loc$Long_,
                     start.date=rep(2000,nrow(xy.loc)), 
                     end.date=rep(2014,nrow(xy.loc)), 
                     ID=1:nrow(xy.loc))

# We need to figure out the name of the product, you can use GetProducts()
GetProducts()
# and check out the data bands that are in the product
GetBands(Product="MOD16A2")

#Create directory in which to save data
if (!dir.exists("MODIS/SantaLucia")) {
  dir.create("MODIS/SantaLucia")
}

# check if directory is empty

# # Now download data using MODISSubsets (This can take very long)
MODISSubsets(LoadDat=coords, Product = "MOD16A2",
             Bands=c("ET_1km","ET_QC_1km"), StartDate=T, Size=c(0,0),
             SaveDir="MODIS")
```

