---
title: "Untitled"
author: "Willem Vervoort"
date: "`r format(Sys.Date(),'%d-%m-%Y')`"
output: 
    pdf_document:
      fig_width: 7
      fig_height: 6
      fig_caption: true
---

```{r setup, warning=F, message=F, echo=F}
# root dir
knitr::opts_knit$set(root.dir = 
                       "X:/vervoort/research/swat")
library(pander)
library(knitr)
```

# Introduction
This series of documents covers different ways of downloading and managing satellite data:

* MODIS16A ET (collection 5) using the package MODISTools
* MODIS13Q1 NDVI (collection 5) (or any other NDVI product) and conversion to ET using METRIC
* Need to talk to Ignacio about Google earth engine

# Downloading MODIS ET data using MODISTools
The first component is a demonstration of using the package MODISTools in R to download the data. In this case the example will be based on extracting data for the midpoints of the subbasins of a SWAT model for the Cotter catchment

## how to extract the midpoints  
This is fairly easy in ARCSWAT, the model actually generates the midpoints as part of the watershed delineation and you can access this through the table of data of the shape file.

## Downloading the data
Below is the actual download of the MODIS data, this requires an input file with the coordinates of the midpoints. Also, at this point in time, there is no MODIS16A in collection 6, so we have to rely on the older collection 5.

```{r download}
# Example catchment  Cotter
if(!require(MODISTools)) install.packages("MODISTools")
library(MODISTools)

# read in file with xy locations (or create a data.frame)
xy.loc <- read.csv("Inputdata/subbasins.csv")
# Following the MODIS tools manual
coords <- data.frame(lat=xy.loc$Lat, 
                     long=xy.loc$Lon,
                     start.date=rep(2000,nrow(xy.loc)), 
                     end.date=rep(2014,nrow(xy.loc)), 
                     ID=1:nrow(xy.loc))

# We need to figure out the name of the product, you can use GetProducts()
GetProducts()
# and check out the data bands that are in the product
GetBands(Product="MOD16A2")

#Create directory in which to save data
if (!dir.exists("MODIS")) {
  dir.create("MODIS")
}

# check if directory is empty

# # Now download data using MODISSubsets (This can take very long)
MODISSubsets(LoadDat=coords, Product = "MOD16A2",
             Bands=c("ET_1km","ET_QC_1km"), StartDate=T, Size=c(0,0),
             SaveDir="MODIS")
```

